Are We Reusing Outdated and Licensed Code from Stack Overflow?

### 2017-02-15 cloneoverflow-latex/cloneoverflow.tex
Updated throughout the paper according to Matheus’s comments. Ready for Jens to look at.

\section{Introduction}
%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 6.3 million users. It allows programmers to ask questions and give answers to programming problems. The website has found to be useful for software development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014, Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also valuable for educational purposes~\cite{Nasehi2012}. On Stack Overflow, each conversation contains a question and a list of answer(s).  The answers frequently contain at least one code snippet as a solution to the question asked. The code snippet is usually not written directly on the Stack Overflow website but copied from another location. It can be copied and modified from a code snippet in the question, copied from an answerer's own code or from other locations including open source software (OSS) systems. The process of posting and answering questions on Stack Overflow, which involves copying and pasting source code, can be considered as code cloning.

Code cloning is an activity of reusing source code by copying and pasting. It normally occurs in software development and account from 7\% to 23\% of source code in typical software systems~\cite{Bellon2007}. The benefits and drawbacks of clones are still controversial. Several authors state that clones lead to bug propagations and software maintenance issues~\cite{Kamiya2002}, while some others suggest that in some cases clones are not harmful or even beneficial~\cite{Saini2016,Kapser2006}.

Code cloning can also have side effects such as violating software licenses and software vulnerabilities. Carelessly cloning code from one project to another project with a different license may cause a software licensing violation~\cite{German2009}. This also happens within the context of online Q\&A websites such as Stack Overflow. An et al.~\cite{An2017} showed that 1,279 cloned snippets between Android apps and Stack Overflow potentially violate software licenses. Security is also among the main concerns when code is copied from an online source. Stack Overflow helps developers to solve Android programming problems more quickly than other resources while, at the same time, offers less secure code than books or the official Android documentation~\cite{Acar2016}.

@@

We call code snippets that are copied from software systems to online Q\&A websites (such as Stack Overflow), and vice versa as ``online code clones'' (sometimes only ``clones'' for brevity). There are four ways to create online code clones: (1) code is cloned from a software project to a Q\&A website as an example; (2) code is cloned from a Q\&A website to a software project to obtain a functionality, perform a particular task, or fixing a bug; (3) code is transferred from one software project to another via a Q\&A website; and (4) code is cloned from an external source to both a software project and a Q\&A website. Online code clones can similarly lead to a problem of outdated code, bug propagation, licensing violation, and software vulnerability as classical code clones. Unfortunately, they are more difficult to locate and fix since the search space in online code corpora is larger and no longer confined in a local repository.

@@

A motivating example of outdated online code clones can be found in the a Stack Overflow post regarding how to implement {\small{\texttt{RawComparator}}} in \texttt{hadoop}\footnote{http://stackoverflow.com/questions/22262310}. In \Cref{fig:before-after}, the left hand side shows a code snippet embedded as a part of accepted answer to the question. The snippet shows how {\small{\texttt{hadoop}}} implements the {\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}} class. The code snippet on the right hand side shows another version of the same {\small{\texttt{compare}}} method in {\small{\texttt{WritableComparator}}} class, but at this time extracted from the latest version (as of 26 September 2016) of {\small{\texttt{hadoop}}}. We can obviously see that they are highly similar except a line containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in the latest version after {\small{\verb|key2.readFields(buffer);|}}. The added line is intended for cleaning up the reference in the {\small{\verb|buffer|}} variable. While this change has already been introduced into the {\small{\texttt{compare}}} method in the latest version of {\small{\texttt{hadoop}}}, the code example in Stack Overflow post is still unchanged. This example shows that there can be inconsistencies between online code clones and their originals leading the users of Stack Overflow to reuse outdated code. %The code snippet on Stack Overflow can be outdated.  %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases. 
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

@@

\vspace{0.5ex}%
\noindent\textbf{1.~A manual study of online code clones:} 
We used two clone detection tools to discover 315,786,118 similar code snippet pairs followed by a manual classification of 3,632 candidate clone pairs between Java code snippets obtained from Stack Overflow's accepted answers and 111 Java open source projects from a curated Qualitas corpus~\cite{QualitasCorpus}.

\vspace{0.5ex}%
\noindent\textbf{2.~Investigating the problems of outdated and license-violating online clones on Stack Overflow:} Our study shows that there are at least 668 clones that have been obviously copied from open source projects or external online sources to Stack Overflow as code examples which potentially violate software licenses. In addition, 64 out of the 112 online code clones from Qualitas to Stack Overflow are outdated and questionable for being reused.

\vspace{0.5ex}%
\noindent\textbf{3.~Online code clone oracle:} The 3,632 manually investigated and validated online clone pairs are available for download on the study website\footnote{https://to.be.disclosed.after.review} and can be used as a clone oracle.

\section{Empirical Study}
We performed an empirical study of online code clones between Stack Overflow and 111 Java open source projects to answer the following research questions: \\ 
\textbf{RQ1 (online code clones):} \textit{To what extent  is source code cloned between Stack Overflow and open source projects?} We quantitatively measured the number of online code clones between Stack Overflow and open source projects to understand the scale of the problem. \newline
\textbf{RQ2 (Patterns of online code clones):} \textit{Why do online code clones occur?} We categorised online clones into seven categories according to our online code cloning pattern scheme. This gave us more insights to why online code clones are created.
\newline
\textbf{RQ3 (Outdated online code clones):} \textit{Are online code clones up-to-date compared to their counterparts in the original projects?} We were interested in the outdated Stack Overflow code examples since millions of users are potentially reusing them. \newline
\textbf{RQ4 (Software licensing violation):} \textit{Do licensing conflicts occur between Stack Overflow clones and their originals?} We investigated whether online code clones can be legally harmful to software reusing them.

@@

Agreed clone pairs are clone pairs that pass Bellon's \textit{good} or \textit{ok} criteria, and therefore are selected for manual classification. Similar to the original study, we select a threshold \textit{p} of 0.7 for both \textit{good} and \textit{ok} pairs~\cite{Bellon2007}. %The number of projects processed by the clone detection tools are listed in \Cref{tab:projects_missing}. 
We encountered NiCad failures with a few Qualitas projects.~$N_D$ could not detect clones in {\small\texttt{hibernate}} due to clustering errors. $N_E$ generated renaming errors for 5 projects including {\small\texttt{vuze}}, {\small\texttt{hibernate}}, {\small\texttt{myfaces}}, {\small\texttt{netbeans}}, and \texttt{spring}. We had contacted the creator of NiCad regarding the issues. They identified the issues as problems of long file paths and TXL grammars which will be fixed in the next NiCad releases. As a result, we did not have NiCad clones from these projects in the agreed clone pairs. Simian clones of the same projects also disappeared from the agreed clone pairs since they could not find matches.

@@

The benefit of the study is two-folded. We evaluated Kapser's cloning patterns and also observed the common cloning patterns from the 697 clone pairs. After the study, we adopted one of Kapser's cloning patterns, boiler-plate code and API/library protocol, and defined 6 new cloning patterns . The seven patterns include QS, SQ, UN, ES, BP, IN, and FC as presented in \Cref{tab:classification_scheme}. Pattern QS (\textbf{Q}ualitas to \textbf{S}tack Overflow) represents clones that has evidence to be copied from Qualitas to Stack Overflow. The evidence of copying can be found in comments in the Qualitas source code or in the Stack Overflow post's contents. Pattern SQ (\textbf{S}tack Overflow to \textbf{Q}ualitas) is cloning, with an evidence, in the opposite direction to QS. Pattern UD (\textbf{U}nknown \textbf{D}irection) is cloning that creates exactly identical or highly similar clones between Qualitas and Stack Overflow but where we could not find any attribution of copying. %The UD clones are also unique enough to not be a boiler-plate code. 
Pattern ES (\textbf{Ex}ternal Sources) is cloning that has evidence of copying from a single or multiple external sources. Pattern BP (\textbf{B}oiler-\textbf{P}late) represents clones containing boiler-plate code of {\small\verb|equals()|} methods, getters/setters, or IDE-generated code such as GUI components. Pattern IN (\textbf{In}heritance/Interface) is cloning by inheritance of the same super class or implementation of the same interface. These two activities usually result in similar overriding methods. The last pattern, AC (\textbf{A}ccidental \textbf{C}lones), represents accidentally similar clone pairs. These are mainly false positive clones from the clone detectors such as similar {\small\texttt{try-catch}} statements or similar code after normalisation.

@@

To search for outdated online code clones, we focused on the 112 QS clone pairs that were cloned from Qualitas to Stack Overflow and compared them with their latest versions. We downloaded the latest version of the Qualitas projects from their repositories on 26 September 2016. For each QS online clone pair, we use the clone from Qualitas a a proxy. We searched for its latest version by the file name and located the cloned region in the file based on the method name. For Simian clones where clones may span across several methods, we scanned upwards until we find a method name ahead of the clone region. We then compared the Stack Overflow snippet to its latest version line-by-line to find if any change has been made to the source code. We also made sure that the changes did not come from the modifications made to the Stack Overflow snippets by the posters but from the updates in the projects themselves. When we found inconsistent lines between the two versions, we used {\small\texttt{git blame}} to see who modified those lines of code and noted down the timestamps and other details.

@@

\textbf{For RQ4, we found 357 code snippets on Stack Overflow that violate the license of their original software. Majority of them do not contain licensing statements after copied to Stack Overflow.}

@@

Kapser et al.~studied clones in Linux file systems and create 11 patterns of code cloning based on four groups: Forking, Templating, Customisation and Exact match~\cite{Kapser2003,Kapser2008}. Our study adopted on of the patterns into our online code cloning patterns.

\textbf{Clone agreement} is useful in when a clone oracle is absent. Since clone detectors are different in their detection approaches, they may behave differently and report different clones even on the same data set. Some researchers exploit these different behaviours of clone detectors by finding their agreement and obtain highly-confident clones~\cite{Bellon2007,Wang2013}. Using the same data set, clone pairs that are agreed by several tools (or several code similarity measurement techniques) are more assured to be true clones than the ones reported by only a single tool~\cite{Wang2013,cr2016ssbse,Funaro2010}. There are studies that report sensitivities of the tools' configurations to the results~\cite{Wang2013,cr2016ssbse} showing that searching for an optimal configuration for every new data set is preferred.

>>>>

\section{Introduction}
%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 6.3 million users. It allows programmers to ask questions and give answers to programming problems. The website has found to be useful for software development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014, Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also valuable for educational purposes~\cite{Nasehi2012}. On Stack Overflow, each conversation contains a question and a list of answer(s).  The answers frequently contain at least one code snippet as a solution to the question asked. We found that the code snippets are usually not written directly on the Stack Overflow website but copied from another location. It can be copied and modified from a code snippet in the question, copied from an answerer's own code or from other locations including open source software (OSS) systems. The process of posting and answering questions on Stack Overflow, which involves reusing of source code, can be considered as code cloning.

Code cloning is an activity of reusing source code by copying and pasting. It normally occurs in software development and account from 7\% to 23\% of source code in typical software systems~\cite{Bellon2007}. The benefits and drawbacks of clones are still controversial. Several authors state that clones lead to bug propagations and software maintenance issues~\cite{Kamiya2002}, while some others suggest that clones are not harmful and can even be beneficial~\cite{Saini2016,Kapser2006}.

Code cloning can also have side effects such as violating software licenses and introducing software vulnerabilities. Carelessly cloning code from one project to another project with a different license may cause a software licensing violation~\cite{German2009}. This also happens within the context of online Q\&A websites such as Stack Overflow. An et al.~\cite{An2017} showed that 1,279 cloned snippets between Android apps and Stack Overflow potentially violate software licenses. Security is also among the main concerns when code is copied from an online source. Stack Overflow helps developers to solve Android programming problems more quickly than other resources while, at the same time, offers less secure code than books or the official Android documentation~\cite{Acar2016}.

@@

We call code snippets that are copied from software systems to online Q\&A websites (such as Stack Overflow), and vice versa as ``online code clones'' (sometimes only ``clones'' for brevity). There are four ways to create online code clones: (1) code is cloned from a software project to a Q\&A website as an example; (2) code is cloned from a Q\&A website to a software project to obtain a functionality, perform a particular task, or fixing a bug; and (3) code is transferred from one software project to another via a Q\&A website; Online code clones can similarly lead to a problem of outdated code, bug propagation, licensing violation, and software vulnerability as classical code clones. Unfortunately, they are more difficult to locate and fix since the search space in online code corpora is larger and no longer confined in a local repository.

@@

A motivating example of outdated online code clones can be found in the a Stack Overflow post regarding how to implement {\small{\texttt{RawComparator}}} in \texttt{hadoop}\footnote{http://stackoverflow.com/questions/22262310}. In \Cref{fig:before-after}, the left hand side shows a code snippet embedded as a part of accepted answer to the question. The snippet shows how {\small{\texttt{hadoop}}} implements the {\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}} class. The code snippet on the right hand side shows another version of the same {\small{\texttt{compare}}} method in {\small{\texttt{WritableComparator}}} class, but at this time extracted from the latest version (as of 26 September 2016) of {\small{\texttt{hadoop}}}. We can obviously see that they are highly similar except a line containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in the latest version. The added line is intended for cleaning up the reference in the {\small{\verb|buffer|}} variable. While this change has already been introduced into the {\small{\texttt{compare}}} method in the latest version of {\small{\texttt{hadoop}}}, the code example in Stack Overflow post is still unchanged. This example shows that there can be inconsistencies between online code clones and their originals leading the users of Stack Overflow to reuse outdated code. %The code snippet on Stack Overflow can be outdated.  %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases. 
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

@@

\vspace{0.5ex}%
\noindent\textbf{1.~A manual study of online code clones:} 
We used two clone detection tools to discover 315,786,118 similar code snippet pairs, followed by a manual classification of 3,632 candidate clone pairs between Java code snippets obtained from Stack Overflow's accepted answers and 111 Java open source projects from a curated Qualitas corpus~\cite{QualitasCorpus}.

\vspace{0.5ex}%
\noindent\textbf{2.~An investigation of outdated and license-violating online clones on Stack Overflow:} Our study shows that there are at least 668 clones that have been obviously copied from open source projects or external online sources to Stack Overflow as code examples which potentially violate software licenses. In addition, 64 out of the 112 online code clones from Qualitas to Stack Overflow are outdated and questionable for being reused.

\vspace{0.5ex}%
\noindent\textbf{3.~An online code clone oracle:} The 3,632 manually investigated and validated online clone pairs are available for download on the study website\footnote{https://to.be.disclosed.after.review} and can be used as a clone oracle.

\section{Empirical Study}
We performed an empirical study of online code clones between Stack Overflow and 111 Java open source projects to answer the following research questions: \\ 
\textbf{RQ1 (online code clones):} \textit{To what extent  is source code cloned between Stack Overflow and open source projects?} We quantitatively measured the number of online code clones between Stack Overflow and open source projects to understand the scale of the problem. \newline
\textbf{RQ2 (Patterns of online code clones):} \textit{Why do online code clones occur?} We categorised online clones into seven categories according to our online code cloning pattern scheme. This gave us more insights to why online code clones are created.
\newline
\textbf{RQ3 (Outdated online code clones):} \textit{Are online code clones up-to-date compared to their counterparts in the original projects?} We were interested in the outdated Stack Overflow code examples since millions of users are potentially reusing them. \newline
\textbf{RQ4 (Software licensing violation):} \textit{Do licensing conflicts occur between Stack Overflow clones and their originals?} We investigated whether online code clones can be legally harmful to software developers reusing them.

@@

Agreed clone pairs are clone pairs that pass Bellon's \textit{good} or \textit{ok} criteria, and therefore are selected for manual classification. Similar to the original study, we select a threshold \textit{p} of 0.7 for both \textit{good} and \textit{ok} pairs~\cite{Bellon2007}. %The number of projects processed by the clone detection tools are listed in \Cref{tab:projects_missing}. 
We encountered NiCad failures with a few Qualitas projects.~$N_D$ could not detect clones in {\small\texttt{hibernate}} due to clustering errors. $N_E$ generated renaming errors for 5 projects including {\small\texttt{vuze}}, {\small\texttt{hibernate}}, {\small\texttt{myfaces}}, {\small\texttt{netbeans}}, and \texttt{spring}. We had contacted the creator of NiCad regarding the issues. They identified these issues as problems of long file paths and TXL grammars which will be fixed in the next NiCad releases. As a result, we did not have NiCad clones from these projects in the agreed clone pairs. Simian clones of the same projects also disappeared from the agreed clone pairs since they could not find matches.

@@

The benefit of the study is two-folded. We evaluated Kapser's cloning patterns and also observed the common cloning patterns from the 697 clone pairs. After the study, we adopted one of Kapser's cloning patterns, boiler-plate code and API/library protocol, and defined 6 new cloning patterns . The seven patterns include QS, SQ, UN, ES, BP, IN, and FC as presented in \Cref{tab:classification_scheme}. Pattern QS (\textbf{Q}ualitas to \textbf{S}tack Overflow) represents clones that has evidence to be copied from Qualitas to Stack Overflow. The evidence of copying can be found in comments in the Qualitas source code or in the Stack Overflow post's contents. Pattern SQ (\textbf{S}tack Overflow to \textbf{Q}ualitas) is cloning, with an evidence, in the opposite direction to QS. Pattern UD (\textbf{U}nknown \textbf{D}irection) is cloning that creates exactly identical or highly similar clones, between Qualitas and Stack Overflow but where we could not find any attribution of copying. %The UD clones are also unique enough to not be a boiler-plate code. 
Pattern ES (\textbf{Ex}ternal Sources) is cloning that has evidence of copying from a single or multiple external sources. Pattern BP (\textbf{B}oiler-\textbf{P}late) represents clones containing boiler-plate code of {\small\verb|equals()|} methods, getters/setters, or IDE-generated code such as GUI components. Pattern IN (\textbf{In}heritance/Interface) is cloning by inheritance of the same super class or implementation of the same interface. These two activities usually result in similar overriding methods. The last pattern, AC (\textbf{A}ccidental \textbf{C}lones), represents accidentally similar clone pairs. These are mainly false positive clones from the clone detectors such as similar {\small\texttt{try-catch}} statements or similar code after normalisation.

@@

To search for outdated online code clones, we focused on the 112 QS clone pairs that were cloned from Qualitas to Stack Overflow and compared them with their latest versions. We downloaded the latest version of the Qualitas projects from their repositories on 26 September 2016. For each QS online clone pair, we use the clone from Qualitas a a proxy. We searched for its latest version by the file name and located the cloned region in the file based on the method name. For Simian clones, where clones may span across several methods, we scanned upwards until we find a method name ahead of the clone region. We then compared the Stack Overflow snippet to its latest version line-by-line to find if any change has been made to the source code. We also made sure that the changes did not come from the modifications made to the Stack Overflow snippets by the posters but from the updates in the projects themselves. When we found inconsistent lines between the two versions, we used {\small\texttt{git blame}} to see who modified those lines of code and noted down the timestamps and other details.

@@

\textbf{For RQ4, we found 357 code snippets on Stack Overflow that violate the license of their original software. The majority of them do not contain licensing statements after copied to Stack Overflow.}

@@

Kapser et al.~studied clones in Linux file systems and create 11 patterns of code cloning based on four groups: Forking, Templating, Customisation and Exact match~\cite{Kapser2003,Kapser2008}. Our study adopted one of the patterns into our online code cloning patterns.

\textbf{Clone agreement} is useful when a clone oracle is absent. Since clone detectors are different in their detection approaches, they may behave differently and report different clones even on the same data set. Some researchers exploit these different behaviours of clone detectors by finding their agreement and obtain highly-confident clones~\cite{Bellon2007,Wang2013}. Using the same data set, clone pairs that are agreed by several tools (or several code similarity measurement techniques) are more assured to be true clones than the ones reported by only a single tool~\cite{Wang2013,cr2016ssbse,Funaro2010}. There are studies that report sensitivities of the tools' configurations to the results~\cite{Wang2013,cr2016ssbse}, showing that searching for an optimal configuration for every new data set is preferred.
